Working Title - CUT Research Paper

### 2019-04-16 CUT/paper.tex
A first draft

\usepackage{hyperref}
\usepackage{apacite}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\doublespace
\title{Working Title - CUT Research Paper}
\author{W. Justin Toth}
\begin{document}
\maketitle
\begin{abstract}
Research Question: ``How does assessment format impact student learning outcomes, and instructor measurement thereof''.

New Research Question: ``Can innovation assessment formats improve student learning outcomes, and an instructors capacity to measure their achievement.''

\textit{With a particular focus on mathematics, stem}
\end{abstract}
\tableofcontents
\section{Introduction}

@@

@@

In an Introduction to Statistics course~\cite{block2012discussion} sought to compare the use of Closed Book, Open Book, and Cheatsheet exams. Over the course of four different semesters, Block varied the exam format and measured student performance and perceptions of the course. In Fall $2006$, the exam format was a traditional Closed Book exam. In Spring $2007$, it was an Open Book exam, and again in Fall $2007$. The difference being, in Fall $2007$, the instructor emphasized to the students that the Open Book exam would be more challenging than a corresponding Closed Book exam and they shoud prepare accordingly using by trying for a deeper understanding of the material. Finally, in Spring $2008$, the Cheatsheet format was tried. Students were allowed to put as much information as they could fit on a typically index card and bring that into the exam with them. Their results showed higher performance on the Closed Book and Cheatsheet exams, and lower performance on the Open Book exams. The students preferred Open Book exams, except when they were told that they would be more difficult exams. The Cheatsheet exam reigned supreme, having comparable student enjoyment to the first semester of Open Book exams, and strong student preparation for the exams in line with what was seen for Closed Book exams. The cheatsheet itself encourages good study habits among students, forcing them to summarize and prioritze content, deciding what is truly important to write down.

@@

@@

\subsection{Collaborative Exams}\label{subsec:collab}
\paragraph{Group}
\paragraph{Two-Stage}

\section{Recommendations}\label{sec:recommendations}
content
\subsection{Designing Exams to Encourage Good Learning Habits}
\subsection{Assessing a Variety of Cognitive Processes}
\subsection{Encouraging Genuine Engagement}
\nocite{*}
\bibliographystyle{apacite}
\bibliography{references.bib}

>>>>

\usepackage{hyperref}
\usepackage{apacite}
\usepackage{setspace}
\usepackage[margin=1in]{geometry}
\doublespace
\title{Working Title - CUT Research Paper}
\author{W. Justin Toth}
\begin{document}
\maketitle
\begin{abstract}
We study of question of how assessment formats impact student study behaviours, learning outcomes, and perception of the courses they take. We particularly focus on Mathematics and Science courses, but the findings in this survey are general enough to be widely applicable. We study Multiple Choice versus Constructed Response, how much Reference Material to allow in the exam room, Oral versus Written Exams, and the Two Stage method for injecting collaborative learning into exams.
\end{abstract}
\tableofcontents
\section{Introduction}

@@

Two common questions when it comes to Written Exams are what sort of question format should be used, and should students be allowed to have access to the textbook during the exam? In this Section we draw from the education research literature to find answers to these questions.

@@

In an Introduction to Statistics course~\cite{block2012discussion} sought to compare the use of Closed Book, Open Book, and Cheatsheet exams. Over the course of four different semesters, Block varied the exam format and measured student performance and perceptions of the course. In Fall $2006$, the exam format was a traditional Closed Book exam. In Spring $2007$, it was an Open Book exam, and again in Fall $2007$. The difference being, in Fall $2007$, the instructor emphasized to the students that the Open Book exam would be more challenging than a corresponding Closed Book exam and they shoud prepare accordingly using by trying for a deeper understanding of the material. Finally, in Spring $2008$, the Cheatsheet format was tried. Students were allowed to put as much information as they could fit on a typically index card and bring that into the exam with them. Their results showed higher performance on the Closed Book and Cheatsheet exams, and lower performance on the Open Book exams. The students preferred Open Book exams, except when they were told that they would be more difficult exams. The Cheatsheet exam reigned supreme, having comparable student enjoyment to the first semester of Open Book exams, and strong student preparation for the exams in line with what was seen for Closed Book exams. The cheatsheet itself encourages good study habits among students, forcing them to summarize and prioritize content, deciding what is truly important to write down.

@@

Getting students to speak up in the classroom is a challenge in its own right. In this Section we consider exam formats which involve students communicating verbally with the instructor and each other. We will find that these approaches break from the norm in North American institutions, but provide many benefits to the students and instructors who can overcome the new challenges they bring.

@@

\subsection{Two-Stage Exams}\label{subsec:collab}
In the previous Subsection (Subsection \ref{subsec:oral}) we briefly discussed student collaboration in the context of an Oral Exam, highlighting the unique challenges and opportunities associated with that exam format. We would now like to discuss an innovative exam format which merges peer-to-peer collaborative discussions with a traditional Written Exam. This is the so-called Two Stage Exam format.
In advance of a Two Stage Exam, students agree on (or are assigned) a team of three to four students who they will collaborate with during the exam period. This already has the advantage of giving students a pre-made study group for exam preparation. The actual exam period is divided into two stages. During the first stage (approximately $2/3$ to $3/4$ of the exam period), students individually write a traditional Written Exam. At the end of the first stage, their exam papers are collected as normal. For the second stage (the remaining time of the exam period), students join up in their pre-arranged groups and each group is given a single copy of the same test they wrote during the first stage. The students are then tasked with coming to a consensus submission for the second stage exam paper. Notice that since the problems are repeats of the first stage problems, they take considerably less time to write their solutions the second time around.
At UBC \cite{rieger2014examinations} tested this Two Stage Exam format in their introductory physics class of $178$ students. The high pressure environment of the first stage of the exam primes students and heavily invests them in finding the answers to the problems. This carries over to the second stage, where students are well prepared to discuss their solution with the group. These discussions yield immediate feedback on student work from their peers. Struggling students receive clarification on problems they did not understand, and stronger students cement their understanding by explaining solutions to their peers. The main reason that Two Stage Exams are great way to get these benefits over regular in class collaborative learning is that they have a much higher engagement rate, nearly $100$ percent according to Rieger and Heiner. They surveyed student opinions of their implementation of Two Stage Exams and found that $76$ percent of students had a generally positive opinion of the format. In particular, the majority of students expressed surprise at how effective the exam method was for them learning from their mistakes.
Students did express some concerns with the Two Stage exam format of~\cite{rieger2014examinations}. Some found it emotionally distressing to immediately be made aware of their exam mistakes. Others expressed concern about weak students unfairly gaining marks from their strong group-mates during the second stage. Finally some students did not appreciate weak or unprepared teammates being unhelpful during the second stage of the exam.

\section{Recommendations}\label{sec:recommendations}
In this paper we discussed some basic decisions instructors need to make about their exam formats and how they can impact the learning outcomes of students in their courses. We have shown that exam format is tremendously impactful on students' experience of the course.
We saw that while Multiple Choice exams are imminently pragmatic, they encourage shallow learning of course material, and are detrimental to long term retention when compared to Constructed Response exam problems. We recommend that instructors relegate Multiple Choice to assessing the lowest levels of Bloom's taxonomy, and even then to only use them in very large classes where the practical benefits outweigh the potential drawbacks.
We also considered the question of allowing students to have access to Reference Material during an exam. We found that Open Book exams do not promote appreciably higher retention of material over Closed Book exams, and can even lead to poor studying practices, like under-preparing for the exam or not employing deep-learning study strategies. The effect of Open Book exams on anxiety seems to be generally positive, reducing test anxiety among students compared to Closed Book exams. But that effect can be lost if you tell students their Open Book exam will be harder than the analogous Closed Book exam. If test anxiety is a major concern for you, then Open Book exams can help to solve that problem. The exam format which seems to capture the best of both worlds is the Cheatsheet exam format. It encourages students to employ similar deep-learning strategies to a Closed Book exam. In fact, it naturally pushes students to prioritize material and assess what is important since they have limited space to record their notes. It also has comparable test anxiety reducing properties to Open Book exams. We highly recommend instructors consider this format for their Written exams. When implementing Cheatsheet exams, ensure to have a standard format for the cheatsheet. Consider limiting both note size and font size, as some students will attempt to squish the entire course on the note somewhat defeating the purpose.
In Section \ref{sec:communication} we studied the potential for Oral Exam formats and Collaboration. We found that Oral Exams provide many impressive benefits, in particular for Mathematics classes. They increase student confidence, allow for immediate corrective feedback and clarification of student understanding, and even allow the instructor to tailor the exam to each individual student based on their level of capability. Perhaps most importantly of all though, there appears to be evidence that they test a different set of aptitudes than Written Exams. Both the skills which are amenable to Oral Exams, and those amenable to Written Exams are desirable skills which should be rewarded and encouraged in the evaluation process.
We strongly recommend that, when practical, instructors incorporate an Oral Exam component into their course. This ``when practical" caveat exists solely because it can be very time-intensive to implement Oral Exams, espcially if your class is large relative to your Teaching Assistant support. The only other reason to avoid Oral Exams may be that your course content is almost entirely mechanical in nature. Courses with little emphasis on conceptual understanding will not see much benefit from Oral Exams. But if those caveats do not exist for your course, or you are not dissuaded by them then Oral Exams can greatly benefit your students. There are a couple tips gathered from the articles we surveyed on Oral Exams which may be of help to instructors. If you have multiple graders for Oral Exams, have some ``practice'' Oral Exams all your Teaching Assistants grade at the start of the term, and discuss the grades as a group to make sure you are all callibrated and equally critical in your assessments. Do not schedule all the Oral Exams on the same day. Spread them out as appointments over a few days, or a week, otherwise the exercise will be mentally fatiguing for the instructor and the quality of exams occuring later in the day will suffer. Some students may freeze out of nervousness, especially if it's their first time giving an Oral Exam. If you know this is a new experience for your class, have some leniency and consider allowing students who break down in this way to reschedule another attempt. Finally, have a rubric which you are grading by present and write on it during the Oral Exam. Do not try to remember and fill it out after the exam. 
Lastly, in Subsection \ref{subsec:collab} we presented an innovative approach to injecting collaborative learning into Written Exams: the Two Stage Exam format. For instructors who have collaborative learning as a high priority goal, this is a relatively easy to implement change to your exam format which can help achieve that goal. It has been found that Two Stage Exams encourage dramatically higher student engagement in the collaborative activity that in-class collaborations. These exams go a long way to to encouraging students to seek feedback on their solutions, identifying missing elements in their understanding, and learn how to correct their mistakes. As with any group work task, poor teamwork skills can lead to drawbacks. To mitigate these, consider a policy where student's group stage mark will not be lower than their individual first stage grade. Similarly, try to make the group work portion just valuable enough that students have an incentive to give an earnest effort, but no higher than that. 
Whether we like it or not, grades are central to the University student experience and are a major motivator for student behaviour. Thus it is not surprising that the way we design our assessments has an impact on how students study, the skills they develop, and ultimately the learning outcomes they achieve in our courses. Hence we should be cognizant of the link between exam format and learning outcomes, and adjust accordingly. In this way we can use student grade motivation to our advantage, designing our assessment formats to funnel students into the positive learning behaviours we all hope they will adopt.
%\subsection{Designing Exams to Encourage Good Learning Habits}
%\subsection{Assessing a Variety of Cognitive Processes}
%\subsection{Encouraging Genuine Engagement}

\bibliographystyle{apacite}
\bibliography{references.bib}
